\documentclass[10pt,letterpaper]{article}
\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[margin=2cm]{geometry}

\usepackage{float}
\usepackage{subfloat}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\newcommand{\dif}[1][]{\mathrm{d} {#1}\,}
\newcommand{\rb}[1]{ \left(  {#1} \right) }
\newcommand{\frb}[1]{ \left(  {#1} \right) }
\newcommand{\norm}[1]{ \left\|  {#1} \right\| }
\newcommand{\jph}{{j+\frac{1}{2}}}
\newcommand{\jmh}{{j-\frac{1}{2}}}
\newcommand{\GodFlx}{G}

% \usepackage[amsmath,thmmarks,standard]{ntheorem}
\newtheoremstyle{break}
{\topsep}
{\topsep}%
{\normalfont}
{}%
{\bfseries}
{}%
{\newline}
{}%
\theoremstyle{break}
\newtheorem{exercise}{Exercise}
\newtheorem*{information}{Information}
\newtheorem{mysolution}{Solution}
% \newenvironment{solution}{\begin{comment}}{\end{comment}}
\newtheorem*{solutioninformation}{Solution Information}

\usepackage{comment}
% Switch between showing and hiding solutions by commenting out either of the following lines
\newenvironment{solution}{\begin{mysolution}}{\end{mysolution}}
% \excludecomment{solution}






\begin{document}

\title{Gudonov's Method}
\date{}

\maketitle








\begin{exercise}
    Discuss qualitatively the derivation of \emph{Godunov's method}; sketch each step in the solution process.
    Which part of the algorithm can make its implementation particularly difficult?
\end{exercise}



\begin{solution}
    Godunov's method can be outlined in two steps.
    Suppose we have an approximation $v^n$ of the solution $u$ at $t_n$.
    (i) Define $\widetilde{u}^{n}(x,t)$ for all $x$ and $t_n<t<t_{n+1}=t_n+k$ as the exact solution to the conservation law, satisfying the initial condition
    \begin{gather} \label{GodIn}
        \widetilde u^n\frb{x,t_n}=v^n_j
        \quad
        x\in(x_{j-1/2},x_{j+1/2})
        \quad\forall j\ .
    \end{gather}
    (ii) Average the resulting function $\widetilde{u}^{n}\frb{x,t_{n+1}}$ over each cell $(x_{j-1/2},x_{j+1/2})$ to obtain the approximation
    \begin{gather} \label{GogAvg}
        v_{j}^{n+1}=\frac{1}{h}\int_{x_{j-1/2}}^{x_{j+1/2}}\tilde{u}^{n}\frb{x,t_{n+1}}\dif x
    \end{gather}
    at $t_{n+1}$.
    Now this procedure can be repeated to advance to the next time-step.

    In Step (i), we need to solve an exact Riemann problem at each cell-interface, over a small time interval $(t_n,t_{n+1})$. Since $\widetilde u^n$ is a solution to the conservation law, \eqref{GogAvg} yields
    \begin{align} \label{Godavg2}
        v_{j}^{n+1} &=\frac{1}{h}\int_{x_{j-1/2}}^{x_{j+1/2}}\tilde{u}^{n}\frb{x,t_{n}}\dif x\\
            &\quad -\frac{1}{h}\rb{
                \int_{t_n}^{t_n+k} f\frb{\tilde{u}^{n}\frb{x_{j-1/2},t}}\dif t
                -\int_{t_n}^{t_n+k} f\frb{\tilde{u}^{n}\frb{x_{j-1/2},t}}\dif t}\ .
    \end{align}
    First notice that
    \begin{gather}
        v^n_j=\frac{1}{h}\int_{x_{j-1/2}}^{x_{j+1/2}}\tilde{u}^{n}\frb{x,t_{n}}\dif x\ ,
    \end{gather}
    since $\widetilde u^n$ satisfies \eqref{GodIn}.
    Now, let us look at the other two integrals in \eqref{Godavg2}.
    Because $\widetilde u^n\frb{\cdot,t_n}$ is piecewise constant, with a discontinuity at each $x_{j-1/2}$, 
    and given that the time step $k$ is sufficiently small, $\widetilde u^n\frb{x_{j-1/2},\cdot}$ is constant. Let
    \begin{gather}
        \tilde{u}^{n}\frb{x_{j-1/2},t}=u^*\frb{v^n_{j-1},v^n_{j}}
        \quad
        t\in(t_n,t_{n+1})\ .
    \end{gather}
    So, if we take
    \begin{gather}
        F(u_l,u_r)=f\frb{u^*(u_l,u_r)}=\frac{1}{k}\int_{t_n}^{t_n+k} f\frb{u^*(u_l,u_r)}\dif t\ ,
    \end{gather}
    then \eqref{GogAvg} becomes
    \begin{gather} \label{GodCons}
        v^{n+1}_j=v^n_j-\frac{k}{h}\Big[ F\frb{v^n_{j+1},v^n_{j}} -F\frb{v^n_{j},v^n_{j-1}}\Big]\ .
    \end{gather}
    This is how the method is applied in practice, provided we can find $u^*$. Unfortunately, evaluating the intermediate can be very expensive, and at times impossible. This motivates the need to construct \textit{approximate Riemann solvers}.
\end{solution}























\begin{exercise}
    Consider the scalar conservation law
    \begin{gather}%
        u_t+f(u)_x=0\ ,
    \end{gather}%
    and initial condition
    \begin{gather}%
        u(x,0)=\begin{cases}
            u_l & x<0\\
            u_r & 0<x
        \end{cases}\ ,
    \end{gather}%
    where the flux $f$ is convex ($f''>0$).
    Godunov's method relies on finding the intermediate state $u^*=u^*(u_l,u_r)$ for which $u(0,t)=u^*$, for $t>0$.
    \begin{enumerate}
    \item
    Show that this intermediate state is given by the following:
    \begin{align*}
        &1.\ \ f'(u_l), f'(u_r)\ge 0 \ \ \Longrightarrow \ \ u^*=u_l \\
        &2.\ \ f'(u_l), f'(u_r)\le 0 \ \ \Longrightarrow \ \ u^*=u_r \\[0.5em]
        &3.\ \ f'(u_l) \ge 0\ge f'(u_r) \ \ \Longrightarrow \ \
            u^*=\begin{cases}
                u_l & s>0 \\
                u_r & s<0
            \end{cases}\ ,
            \qquad s=\frac{f(u_r)-f(u_r)}{u_r-u_l}\\[0.5em]
        &4.\ \ f'(u_l) < 0 < f'(u_r)  \ \ \Longrightarrow \ \ u^*=u_m\ ,
            \qquad\text{where $u_m$ is the solution to $f'\frb{u_m}=0$.}
    \end{align*}

    \item
    Use (a) to show that Godunov's flux is given by
    \begin{gather} \label{GodFlx}
        F(u_l,u_r)=\begin{cases}
                \displaystyle
                \min_{u_l\le u\le u_r} f(u) & u_l\le u_r \\[1em]
                \displaystyle
                \max_{u_r\le u\le u_l} f(u) & u_l> u_r 
            \end{cases}\ .
    \end{gather}%
    \item
    Show that Godunov's flux \eqref{GodFlx} is monotone.
    \end{enumerate}
\end{exercise}


\begin{solution}
    Consider the scalar conservation law
    \begin{gather}
        u_t+f(u)_x=0\ ,
    \end{gather}
    and initial condition
    \begin{align}
        u(x,0)=\begin{cases}
            u_l & x<0\\
            u_r & 0<x
        \end{cases}\ ,
    \end{align}
    where the flux $f$ is convex ($f''>0$).
    Godunov's method relies on finding the intermediate state $u^*=u^*(u_l,u_r)$ for which $u(0,t)=u^*$, for $t>0$.
    \begin{enumerate}
        \item
        We show that $u^*$ is given by
        \begin{align*}
            &1.\ \ f'(u_l), f'(u_r)\ge 0 \ \ \Longrightarrow \ \ u^*=u_l \\
            &2.\ \ f'(u_l), f'(u_r)\le 0 \ \ \Longrightarrow \ \ u^*=u_r \\[0.5em]
            &3.\ \ f'(u_l) \ge 0\ge f'(u_r) \ \ \Longrightarrow \ \
                u^*=\begin{cases}
                    u_l & s>0 \\
                    u_r & s<0
                \end{cases}\ ,
                \qquad s=\frac{f(u_r)-f(u_r)}{u_r-u_l}\\[0.5em]
            &4.\ \ f'(u_l) < 0 < f'(u_r)  \ \ \Longrightarrow \ \ u^*=u_m\ ,
                \qquad\text{where $u_m$ is the solution to $f'\frb{u_m}=0$.}
        \end{align*}
        Note that since $f$ is strictly convex, the Jacobian $f^\prime$ is a strictly increasing function.

        Suppose $f'(u_l), f'(u_r)\ge 0$.
        If $u_l>u_r$, the entropy solution is a shock moving at speed given by the RH condition
        \begin{gather} \label{RHJumpCond}
            s=\frac{f(u_l)-f(u_r)}{u_l-u_r}\,
        \end{gather}
        and $f'(u_l) > s > f'(u_r)$ according to the entropy condition.
        This implies that the shock speed is positive, and thus we have $u^*=u_l$. 
        If $u_l\le u_r$, the entropy solution is a rarefaction wave.
        Since $f'(u_l)>0$, the left front of the wave moves to the right, and thus $u^*=u_l$.

        Next suppose $f'(u_l), f'(u_r)\le 0$.
        By similar arguments we get $u^*=u_r$.

        If $f'(u_l) \ge 0\ge f'(u_r)$, the entropy solution is a shock, and the intermediate state $u^*$ is determined by the sign of the shock speed \eqref{RHJumpCond}.

        Finally suppose $f'(u_l) < 0 < f'(u_r)$.
        Then, the entropy solution is a rarefaction wave.
        This time $x=0$ falls inside the rarefaction fan.
        As we have seen in a previous exercise (see exercise set 3), the rarefaction solution is given by $u(x,t)=w\frb{x/t}$, where $w$ is the solution to $f'\frb{w\frb{\xi}}=\xi$.
        Since we are interested in the value of $u$ at $x=0$, we have $u^*=w\frb{0}$.

        \item
        Before we begin, let as remark that since $f$ is strictly convex, the its maximum on a given closed interval $[u_1,u_2]$ is achieved its maximum at an end point of the interval. Furthermore, there exists a unique $\theta$ where $f^\prime$ vanishes. This point corresponds to the global minima of $f$. If $\theta \in (u_1,u_2)$, then f achieves its minimum at $\theta$, else at an end point of the interval.
        To show that Godunove's flux is given by
        \begin{gather} \label{GodFlx}
            F(u_l,u_r)=\begin{cases}
                    \displaystyle
                    \min_{u_l\le u\le u_r} f(u) & u_l\le u_r \\[1em]
                    \displaystyle
                    \max_{u_r\le u\le u_l} f(u) & u_l> u_r 
                \end{cases}\ .
        \end{gather}
        we look at the different possible cases.

        If $u_l>u_r$, the entropy solution is a shock.
        In this case, $u^*$ is determined by the sign of $s$.
        If $f(u_l)>f(u_r)$, then $u^*=u_l$, and if $f(u_l)<f(u_r)$, then $u^*=u_r$.
        Either way, $f\frb{u^*}$ is the maximum of $f$ in $[u_r,u_l]$.

        If $u_l\le u_r$, the entropy solution is a rarefaction wave.
        This is possible only when 1,2 or 4 are valid.
        If 1 is valid, $f$ is increasing in $[u_l,u_r]$, and $f(u_l)$ is the minimum of $f$ in this interval.
        Similarly, if 2 is valid, $f$ is decreasing in $[u_l,u_r]$, and $f(u_r)$ is the minimum of $f$ in this interval.
        If 4 is valid, $f$ achieves its minimum at the internal point $u_m$, where $f'\frb{u_m}=0$.


        \item
        To show that Goduno's flux, given by \eqref{GodFlx}, is monotone, we show that it is non-decreasing in its first argument and non-increasing in its second argument.
        If $u_l<u_r$ and $\epsilon>0$ is small enough, then
        \begin{gather}
            F\frb{u_l+\epsilon,u_r}=\min_{u_l+\epsilon\le u\le u_r} f(u) \ge \min_{u_l\le u\le u_r} f(u)
                =F(u_l,u_r)\ ,
        \end{gather}
        and if $u_l\ge u_r$, then
        \begin{gather}
            F\frb{u_l+\epsilon,u_r}=\max_{u_r\le u\le u_l+\epsilon} f(u) \ge \max_{u_r\le u\le u_l} f(u)
                =F(u_l,u_r)\ .
        \end{gather}
        Similarly, one can show that $F$ is non-increasing in its second argument.
    \end{enumerate}
\end{solution}
























\begin{exercise}
    The purpose of this exercise is to illustrate the \emph{Lax-Wendroff Theorem}.
    This theorem states that if there exists a sequence $\{(h_l,k_l)\}_l^\infty$ with $k_l = \lambda h_l$ ($\lambda$ is kept constant), such that the corresponding numerical solutions $\left\{ v_{l}\right\} _{l=1}^{\infty}$ obtained by a conservative method converges to some function $u$, then the limit $u$ is a weak solution of the conservation law. Notice that to deduce the conclusion, we assume that $v_{l}$ converges as $l\rightarrow\infty$.
    That is, convergence is not a conclusion of the Lax-Wendroff theorem.
    Also recall that in general weak solutions are not unique, so the theorem does not guarantee the limit is the correct entropy solution. 

    Consider a conservative method 
    \begin{gather}%
        v_{j}^{n+1}=v_{j}^{n} -\frac{k}{h}\left(
            F\frb{v_{j}^{n},v_{j+1}^{n}}-F\frb{v_{j-1}^{n},v_{j}^{n}}
            \right)
    \end{gather}%
    where the numerical flux F is given by 
    \begin{gather}%
        F\frb{v,w}=\begin{cases}
            f(v) & \frac{f(v)-f\frb{w}}{v-w}\geq0\\[0.5em]
            f\frb{w} & \frac{f(v)-f\frb{w}}{v-w}<0
        \end{cases}\ .
    \end{gather}%
    \begin{enumerate}
        \item
        Construct the entropy solution to the following initial value problem
        \begin{gather} \label{IVP}
            u_{t}+\rb{\frac{1}{2}u^{2}}_{x}=0
            \quad
                u(x,0)=\begin{cases}
                -1 & x<1\\
                1 & x>1
            \end{cases}\ .
        \end{gather}%
        %satisfying the initial condition
        %\begin{gather} \label{incond2}
        %	u(x,0)=\begin{cases}
        %		-1 & x<1\\
        %		1 & x>1
        %	\end{cases}\ .
        %\end{gather}%
        \item
        Fix $k/h=0.5$, and implement the above method to \eqref{IVP}, in $x\in(0,2)$, $0<t\le0.25$, with the initial data 
        discretized using cell averages. On the boundaries, set $u(0,t)=-1$, and $u\frb{2,t}=1$.

        \item Run the computations by choosing i) $h_l = \frac{2}{l}$, ii)$h_l = \frac{2}{2 l}$, iii) $h_l = \frac{2}{2 l + 1}$, for $l\in\mathbb{N}$.

        \item
        What can you deduce from your results regarding the each of the three sequences of numerical solutions obtained.
        Explain your results and conclude how they fit with the Lax-Wendroff theorem.
    \end{enumerate}
\end{exercise}

\begin{solution} \label{exLWTNonConvSeq}
    \begin{enumerate}
    \item
    Since Burgers equation is convex and
    \begin{gather} \label{InCond}
        u(x,0)=\begin{cases}
            -1 & x<1\\
            1 & x>1
        \end{cases}\ ,
    \end{gather}
    that is $u_l<u_r$, a shock is not admissible.
    Therefore, the entropy solution must be a rarefaction wave.
    The exact solution to the problem with initial data \eqref{InCond} is given by
    \begin{gather}
        u(x,t)=\begin{cases}
            -1 & x\leq-t+1\\
            \frac{x-1}{t} & -t+1<x\leq t+1\\
            1 & \; x>t+1
        \end{cases}\ .
    \end{gather}

    \item[(b),(c)]
    \addtocounter{enumii}{2}
    See Matlab code attached at the end of this manual. Experiment with different values of $l$.

    \item
    We notice that in case of $k_{l}=\frac{1}{2l}$, for any choice
    of $l\in\mathbb{N}$, we get an approximation to the entropy solution,
    and that the sequence of solutions converges as $l\rightarrow\infty$.
    In the case $k_{l}=\frac{1}{2l+1}$, for any choice of $l\in\mathbb{N}$ we get the entropy violating solution of a stationary discontinuity.
    Therefore, the sequence of solutions obtained by taking $k_{l}=\frac{1}{2l+1}$ converges to the same entropy violating solution as $l\rightarrow\infty$.
    Note that in either case, the converged solution is still a weak solution, which is in accordance with the Lax-Wendroff theorem. However, by taking $k_{l}=\frac{1}{l}$, the numerical solution does not converge to any solution with the current scheme.
    \end{enumerate}
\end{solution}


% \begin{figure}[H] 
%     \center
%     \subfloat[\label{fig:Exc4_2a_U}]{\input{Exc4_2a_U.tikz}}
%     \hspace{10mm}
%     \subfloat[\label{fig:Exc4_2a_C}]{\input{Exc4_2a_C.tikz}}
%     \caption{\label{fig:Exc4_2b} Exercise 1.3b. Initial data $u_0(x)$ Eq. 16 applied to the inviscid burgers equation. The two shocks move away from each other with equal speed and never merge. \textbf{(a)} Initial data. \textbf{(b)} Characteristics. }
% \end{figure}

\newpage
\begin{lstlisting}

% Solution05: Problem 3
% This script was written for EPFL MATH459, Numerical Methods for
% Conservation Laws. The invisvid Burgers equation is solved using the
% Roe flux

clc
clear all
close all

% Initial left and right states for the Riemann Problem
U0 = @(x) -1*(x<1) + 1*(x>1);

FinalTime = 0.25;

% Roe flux
RoeFlux =@(UL,UR) 0.5*UL.^2.*( (UL+UR) >=0) + 0.5*UR.^2.*( (UL+UR) <0);


% Discretization
l = 50:10:100;
%l = 51:10:101;
%l = 50:5:100;

M = length(l);

for i=1:M
    h = 2/l(i);
    k = 0.5*h;
    xf = 0:h:2;
    xc = 0.5*h:h:2-0.5*h;
    N  = length(xc);
    
    % Finding cell-average values of initial condition
    % Since we are looking at first-order schemes, it is enough to
    % approximate the integrals using the mid-point quadrature rule. This
    % essentially means we take the cell-center values as the cell-averages
    U = U0(xc);
    
    time = 0;
    while(time<FinalTime)
        
        if(time+k>FinalTime)
            k = FinalTime-time;
        end
        
        U_ext = [U(1),U,U(end)];
        
        Flux  = RoeFlux(U_ext(1:end-1),U_ext(2:end));
        
        U = U - (k/h)*(Flux(2:end) - Flux(1:end-1));
        
        time = time + k;
    end
    
    figure(1)
    plot(xc,U,'LineWidth',2)
    title(['Solution on a mesh with h = 2/',num2str(l(i))])
    hold all
    pause(0.1)
    
end




\end{lstlisting}
\end{document}
