\documentclass{article}



\usepackage{../auxFiles/ExStyPac}



\input ../auxFiles/mac.tex
\graphicspath{{./figures/}}





%========================================================================
\begin{document}



\Head{12}{Discontinuous Galerkin}










Let $\Om=\Om_x\times\Om_t$, where $\Om_x=\rb{-1,1}$ is the spacial domain and $\Om_t=\rb{0,T}$ is the time domain.

\begin{exerciseList}

%\item
%Consider the scalar problem
%\begin{equation}
%	u_t+au_x=bu
%	\quad
%	x\in\rb{-1,1}
%\end{equation}
%with proper initial conditions and $a$ and $b$ being real constants.
%\textbf{(a)}
%Propose a discontinuous Galerkin method for solving this problem.
%\textbf{(b)}
%Prove that the semi-discrete scheme is stable. 



\item
Consider the scalar PDE
\begin{gather} \label{WEH}
	u_t+au_x=g(x,t)
\end{gather}%
in $\Om$, with proper initial conditions and periodic boundary conditions.
Here $a>0$ is a real constant.
\begin{enumerate}

\item
Write the weak discontinuous Galerkin (DG) formulation for the problem.

\item
Pick an appropriate numerical flux and prove that the semi-discrete scheme is stable.
% (Hint: To prove stability, consider the problem for the error between the computed and the exact solution - known as the error equation).
\end{enumerate}




\item
Consider the system 
\begin{gather}%a
	u_{t}+v_{x} &=0\\
	v_{t}+u_{x} &=0
\end{gather}%a
in $\Om$, subject to periodic boundary conditions.
\begin{enumerate}

\item
Write the weak DG formulation for the problem.

\item
Show that the formulation with the upwind flux is stable.
\end{enumerate}



\item
Consider the ODE system
\begin{gather} \label{odeSys}
	u'=L(u)\ .
\end{gather}%
Suppose there exists a positive constant $k_{FE}$ such that the forward Euler method
$$
	v^{n+1}=v^n+kL\frb{v^n}
$$
with $0< k\le k_{FE}$, applied to \eqref{odeSys} satisfies
\begin{gather} \label{stabEst}
	\|v^{n+1}\|\le\|v^n\|
\end{gather}%
in some norm $\|\cdot\|$.
Show that the numerical approximation $U$ of \eqref{odeSys} obtained by the following 3rd-order Runga-Kutta method
\begin{align*}
	U^{\left(1\right)}= & U^{n}+kL\left(U^{n}\right)\\
	U^{\left(2\right)}= & \frac{3}{4}U^{n}+\frac{1}{4}U^{\left(1\right)}+\frac{1}{4}kL\frb{U^{\left(1\right)}}\\
	U^{n+1}= & \frac{1}{3}U^{n}+\frac{2}{3}U^{\left(2\right)}+\frac{2}{3}kL\frb{U^{\left(2\right)}}
\end{align*}
satisfies \eqref{stabEst}, provided $0< k\le k_{FE}$.


%\item
%Find the \emph{nodal} DG code on the course website.
%\begin{enumerate}[label={\textbf{(\alph*)}}]
%
%\item
%Modify the solver for the scalar advection equation to solve 
%\begin{gather} \label{WEg}
%	u_t+au_x=h(x,t)
%	\quad
%	x\in\rb{0,2}
%\end{gather}%
%with $a>0$ and 
%\begin{gather}%
%	u(0,t)=g\frb{t}\ ,\quad u(x,0)=f(x)\ .
%\end{gather}%
%
%\item
%Verify the convergence rate of the scheme for various $N$ on \eqref{WEg} with $h(x,t)=0$.
%\end{enumerate}
%



\end{exerciseList}

\end{document}
